---
output:
  word_document: default
  html_document: default
---
## Simple Linear Regression and Correlation - Module 2 Assignment 1

###  Nicholas Barrett, 02/02/19


###  Task 1 - Read-in the airquality dataset as a df called "air".


```{r}
library(tidyverse)
library(GGally)
```

```{r}
air <- airquality
```

```{r}
str(air)
```

```{r}
summary(air)
```

a.  All of the variables (Ozone, Solar.R, Wind, Temperature, Month and Day) in the new "air" dataset are represented as numeric.

b.  The "air" dataset contains 153 observations of 6 variables 

c.  The Ozone and Solar R variables as missing data (37 in "Ozone" and 7 in "Solar.R")

d.  Ozone levels is the most likely response variable (Y)

###  Task 2 - Missing data in the two variables: "Ozone" and "Solar.R"  Three options - Delete the rows with missing data, delete the columns with missing data or impute values to replace missing values.  Delete rows with missing data.



```{r}
air2<- air %>% filter(!is.na(Ozone)) %>% filter(!is.na(Solar.R))
```

The new "air2" dataset has 111 rows and 6 columns.

###  Task 3 - Use the code provided to develop a visualization of and to calculate correlation for the combinations of variables in this dataset.  Use "ggcorr" to develop a correlation matrix for the variables. 



```{r}
ggpairs(air2)
ggcorr(air2,label = TRUE)
```

a.  The variable that most strongly correlates with the "Ozone" variable is "Temp".

b.  The variable that least strongly correlates with the "Ozone" variable is "Wind".

###  Task 4 - Plot "Temp" vs. "Ozone" using "ggplot" using appropriate chart typle.  Describe the relationship between the two variables.



```{r}
ggplot(air2, aes(x = Temp, y = Ozone)) + 
  geom_point()
```

Based on the plot above, "Temp" appears to have an increasing linear relationship with "Ozone" with some outliers. As temparatures rise, ozone levels appear to rise.

###  Task 5 - Create a linear regression model (model1) using "Temp" to predict "Ozone".


```{r}
model1<- lm(Ozone ~ Temp, air2)
summary(model1)
```

a.  The above "model1" model for the "air2" appears to be a decent predicative model for determining Ozone from the Temp variables.  The Temp, predictor, variable is significant with a p-value less than 0.05.  The R-squared value is pretty good at 0.4833.

b.  Based on the 95% confidence interval below,the slope co-efficent falls between 1.964787 and 2.913433


```{r}
confint(model1)
```

###  Task 6 - Re-do Task 4 


```{r}
ggplot(air2, aes(x = Temp, y = Ozone)) + geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "red")+
  theme_bw()
```

###  Task 7 - Develop a prediction and prediction interval for "Ozone" when "Temp" is 80.


```{r}
testdata = data.frame(Temp = 80)
predict(model1, newdata = testdata, interval = "predict")
```

The prediction for Ozone when Temp is 80 is 47.48272 with a predicaiton interval between -0.1510188 lower and 95.11646 upper.

###  Task 8 - Plot "Day" vs. "Ozone" using ggplot.  Describe the relationship between the two variables.


```{r}
ggplot(air2, aes(x = Day, y = Ozone)) + geom_point()
```

Considering the plot above, there does not appears to be an apparent or significant linear relationship between "Day" and "Ozone".

###  Task 9 - Create a linear regression model called "model2" using "Day" to predict "Ozone".


```{r}
model2<- lm(Ozone ~ Day, air2)
summary(model2)
```

a.  The above "model2" model for the "air2" appears to be a very poor predicative model for determining Ozone from the Day variables.  The Day, predictor, variable is insignificant with a p-value more than 0.05 (0.957).  The R-squared value is also bad at -0.009147.

b.  Based on the 95% confidence interval below,the slope co-efficent falls between -0.745321 and 0.7056539

```{r}
confint(model2)
```

###  Task 10 - Re-do Task 8 to inlcude the regression line.



```{r}
ggplot(air2, aes(x = Day, y = Ozone)) + geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "red")+
  theme_bw()
```



