---
output:
  word_document: default
  html_document: default
---
## Simple (One Predictor) Linear Regression

Needed libraries
```{r}
library(tidyverse)
library(GGally)
library(ggcorrplot)
```

Read-in the data.  Before doing this make sure that you have placed the CreditData.csv file (downloadable from Canvas) in your projects working directory.

```{r}
credit <- read_csv("CreditData.csv")
```

Examine the structure and summary of the dataset
```{r}
str(credit) #all variables numeric
summary(credit)  #notice 896 missing rows
```

Get rid of missing data rows
```{r}
credit <- credit %>% drop_na() #delete any row with an NA value
str(credit) #check structure after the drop
```

Our Y (response) variable in the dataset is "AnnualCharges".  Let's look at ggparis plot for visualization and correlation.
```{r}
ggpairs(credit)
```

The best variable (by correlation and confirmed by visualization) to predict AnnualCharges appear to be AnnualIncome (correlation = 0.562 and there is an intuitive increase in charges as income increases).

Build a regression model with AnnualIncome to predict AnnualCharges.
```{r}
mod1<-lm(AnnualCharges ~ AnnualIncome, credit) #create linear regression model
summary(mod1) #examine the model
```

Is this a good model?

```{r}
ggplot(credit,aes(x=AnnualIncome,y=AnnualCharges))+geom_point()+geom_smooth(method="lm", color="red")+ theme_bw()
```

```{r}
mod2<-lm(AnnualCharges ~ HouseholdSize, credit) #create linear regression model
summary(mod2) #examine the model
```

```{r}
ggplot(credit,aes(x=HouseholdSize,y=AnnualCharges))+geom_point()+geom_smooth(method="lm", color="red")+ theme_bw()
```

Is this a good model?

R-squared value is pretty poor.  The HouseholdSize variable is significant (p-value is < 0.05)and has an intuitive sign.  Note:  As datasets increase in size, it is very easy for the predictor vairable to be significant.



