---
output:
  word_document: default
  html_document: default
---
##  Module 5 Assignment 1 Neural Networks

###  Nicholas Barrett, 02/23/19

```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(caret)
library(nnet)
```

```{r}
parole = read_csv("parole.csv")
```

```{r}
parole = parole %>% mutate(male = as.factor(male)) %>% 
  mutate(male = fct_recode(male, "Female" = "0", "Male" = "1" ))
parole = parole %>%mutate(race = as.factor(race)) %>%
  mutate(race = fct_recode(race, "Other" = "2", "White" = "1" ))
parole = parole %>% mutate(state = as.factor(state))%>%
  mutate(state = fct_recode(state, "Kentucky"= "2", "Lousiana"= "3", "Virginia"= "4", "Other"= "1" ))
parole = parole %>% mutate(crime = as.factor(crime))%>%
  mutate(crime = fct_recode(crime, "Larceny"= "2", "Drug-Related"= "3", "Driving-Related"= "4", "Other"= "1" ))
parole = parole %>% mutate(multiple.offenses = as.factor(multiple.offenses))%>%
  mutate(multiple.offenses = fct_recode(multiple.offenses, "Incarcerated" = "1", "Other" = "0"  ))
parole = parole %>% mutate(violator = as.factor(violator))%>%
  mutate(violator = fct_recode(violator, "Violated" = "1", "No Violation" = "0"  ))
```

##  Task 1 - Split the data into training and testing sets. Training set should have 70% of the data.  Use a set seed of 12345

```{r}
set.seed(12345)

train.rows = createDataPartition(y = parole$violator, p=0.7, list = FALSE) 
train = parole[train.rows,] 
test = parole[-train.rows,]
```

##  Task 2 - Create a neural network to predict parole violation. Use 12 and a decay rate of 0.1.  Use caret to implement 10-fold k-fold cross-validation.  Use a random number seed of 1234.

```{r}
start_time = Sys.time()
fitControl = trainControl(method = "cv", number = 10)

nnetGrid <- expand.grid(size = 12, decay = 0.1)

set.seed(1234)
nnetTrain = train(violator ~., 
                  train,
                  method = "nnet",
                  tuneGrid = nnetGrid,
                  trControl = fitControl,
                  verbose = FALSE,
                  trace = FALSE)

end_time = Sys.time()
end_time-start_time
```

```{r}
nnetTrain
```


##  Task 3 - Use the model from Task 2 to develop predictions on the training set.  Use caret's confusionMatrix function to evaluate the model for quality.  Comment on the quality.
```{r}
predNetTrain = predict(nnetTrain, train)
```

```{r}
confusionMatrix(predNetTrain, train$violator, positive = "Violated")
```

The nueral network model's accuracy with the size to be 12 and decay to be 0.1 was 88.3% accurate.  After tuning the model and using caret's confusionnMatric, the accuracy improved to 94.7% when applied to the training set.  So, the model is quite good.

##  Task 4 - Create a neural network to predict parole violation.  use a grid to search sizes 1 through 12 (by 1) and decay rates of 0.1 to 0.5 (by 0.1).  use caret to implement 10-fild k-fold cross-validation.  Use a random number seed of 1234.  Suppress the text describing model convergence using the command trace = FALSE.


```{r}
start_time = Sys.time()
fitControl = trainControl(method = "cv",
                          number = 10)

nnetGrid = expand.grid(size = seq(from = 1, to = 12, by = 1),
                       decay = seq(from = 0.1, to = 0.5, by = 0.1))

set.seed(1234)
nnetFitTrain = train(violator ~., 
                  train,
                  method = "nnet",
                  tuneGrid = nnetGrid,
                  trControl = fitControl,
                  verbose = FALSE,
                  trace = FALSE)

end_time = Sys.time()
end_time-start_time
```


##  Task 5 - Use the model from Task 4 to develop predictions on the training set.  Use caret's confusionMatric function to evaluate the quality of the model.  Comment on the quality of the model.

```{r}
nnetFitTrain
```

```{r}
predTrain = predict(nnetFitTrain, train)
```

```{r}
confusionMatrix(predTrain, train$violator, positive = "Violated")
```

The optimal value of the tuning parameters for the training set were a size of 1 and a decay of 0.3.  After tuning, those parameters are different than the training set parameters of a size of 12 and a decay of 0.1.  The confusionMatrix showed an accuracy of 89.6% which isnt quite as good as the model with a 94.7% accuracy.

##  Task 6 - Use the model from Task 2 to develop predictions on the testing set.  Comment on the model quality.

```{r}
start_time = Sys.time()
fitControl = trainControl(method = "cv", number = 10)

nnetGrid <- expand.grid(size = 12, decay = 0.1)

set.seed(1234)
nnetTest = train(violator ~., 
                  test,
                  method = "nnet",
                  tuneGrid = nnetGrid,
                  trControl = fitControl,
                  verbose = FALSE,
                  trace = FALSE)

end_time = Sys.time()
end_time-start_time
```


```{r}
nnetTest
```

```{r}
predNetTest= predict(nnetTest, test)
```

```{r}
confusionMatrix(predNetTest, test$violator, positive = "Violated")
```

The model accuracy with the size to be 12 and decay to be 0.1 was 85.7% accurate.  After tuning the model and using caret's confusionnMatric, the accuracy improved to 95.05% when applied to the training set.  So, the model is quite good.

## Task 7 - Use the model from Task 4 to develop prediction on the testing set.  Use caret's confusionMatrix to assess and comment on the quality of the model.

```{r}
start_time = Sys.time()
fitControl = trainControl(method = "cv",
                          number = 10)

nnetGrid = expand.grid(size = seq(from = 1, to = 12, by = 1),
                       decay = seq(from = 0.1, to = 0.5, by = 0.1))

set.seed(1234)
nnetFitTest = train(violator ~., 
                  test,
                  method = "nnet",
                  tuneGrid = nnetGrid,
                  trControl = fitControl,
                  verbose = FALSE,
                  trace = FALSE)

end_time = Sys.time()
end_time-start_time
```


```{r}
nnetFitTest
```

```{r}
predTest = predict(nnetFitTest, test)
```

```{r}
confusionMatrix(predTest, test$violator, positive = "Violated")
```

The optimal value of the tuning parameters for the training set were a size of 1 and a decay of 0.1.  Those parameters are different than the training set parameters of a size of 12 and a decay of 0.1.  The confusionMatrix showed an accuracy of the tuned testing set model of 89.6% which isnt quite as good as the neural network training model with a 94.7% accuracy.

##  Task 8 - Comment on whether there appears to be overfitting in one or both of your models from Task 2 and 4.

Both models have a very high specificity and accuracy which could potentially mean they are both overfitted.
